Turbo fast no compromise gravitational wave parameter estimation

### 2023-01-11 src/tex/ms.tex
Add citation in introduction

% Brief description of PE
Parameter estimation (PE) is one of the most commonly performed data analysis
tasks in gravitational wave (GW) data analysis. The central question of PE is to
infer the parameters of a particular GW model given the strain data. In the
standard compact binary coalescence scenario, this could mean inferring
intrinsic parameters such the masses and spins of the event, as well as
extrinsic parameters such as sky localization and distance from Earth to the
event. PE is also used in testing general relativity (GR), which the main
question is then to check for deviations in parameters that are related to
non-GR modifications \kw{cite}. PE is a crucial step in GW science, since it translate
characteristics of the strain data into astrophysically relevant quantities that
can be used to constrain astrophysical phenomenon such as theories of binary
evolution \kw{cite}.

% Old codes have trouble handling next gen data
The PE results in the most recent catalog of GW events are powered by several
community-developed PE codes, including \texttt{lalinference}, \texttt{pycbc},
and \texttt{bilby} \cite{Romero-Shaw:2020owr} \kw{cite}. These packages have
been tested by a number of groups and are well regarded as the standard tools.
While the standard tools have passed many robustness tests, they are known to be
computationally intensive. The exact amount of time needed to analyze one event
depends on the length of the event. These community tools take about a week to
analyze shorter events such as binary black hole (BBH) mergers, whereas they can
take up to a month on a computer cluster to analyze longer events such as binary
neutron star (BNS) mergers. In the coming decade, there are planned upgrade for
existing facilities and next-generation detector such as the Einstein Telescope
(ET) \kw{cite} and the Cosmic Explorer (CE) \kw{cite}. These upgrades will
increase the sensitivity of the detectors and allow for the detection of more
events with better signal-to-noise ratio (SNR). The number of events that will
be detected in the coming decade is expected to grow from around a thousand per
year to over a million per year. This will put a significant strain on the
current PE tools, which are not designed to handle this many events.

% Effort going to new tools development.
In order to handle the future catalog, there are efforts from multiple groups
developing tools to speed up the PE process. This includes methods that employ
modern tools such as deep learning-based method pre-trained on a large
collection of waveforms \kw{cite}, as well as methods that simplify the
computational challenge by leveraging our knowledge about GW signals \kw{cite}.
While these methods are promising avenues to tackle standard GW analysis tasks
in the coming decade, particularly about events involving compact binary coalescence
(CBC) in GR, they capitalize assumptions that may not be valid for analysis involving
outliers and beyond-GR analysis.

@@

For a quantitative comparison between results obtained through our code and
other existing tools, we compute the Jensen-Shannon divergence of the
marginalized distribution between our code and \texttt{Bilby}. The
Jensen-Shannon divergence is a symmetric measure of the distance between two
probability distributions, with a value of 0 indicating identical distributions
and a value of $\ln{2} \textrm{nat}$ representing the maximum possible
divergence between two distributions.
\kw{Report JSO}

>>>>

% Brief description of PE
Parameter estimation (PE) is one of the most commonly performed data analysis
tasks in gravitational wave (GW) data analysis. The central question of PE is to
infer the parameters of a particular GW model given the strain data
\cite{Christensen:2022bxb, 2019PASA...36...10T}. In the standard compact binary
coalescence scenario, this could mean inferring intrinsic parameters such the
masses and spins of the event, as well as extrinsic parameters such as sky
localization and distance from Earth to the event. PE is also used in testing
general relativity (GR), which the main question is then to check for deviations
in parameters that are related to non-GR modifications
\cite{LIGOScientific:2016lio,LIGOScientific:2018dkp,LIGOScientific:2021sio}. PE
is a crucial step in GW science, since it translate characteristics of the
strain data into astrophysically relevant quantities that can be used to
constrain astrophysical phenomenon such as theories of binary evolution
\cite{LIGOScientific:2021psn}.

% Old codes have trouble handling next gen data
The PE results in the most recent catalog of GW events are powered by several
community-developed PE codes, including \texttt{lalinference}
\cite{Veitch:2014wba}, \texttt{pycbc} \cite{Biwer:2018osg}, and \texttt{bilby}
\cite{Ashton:2018jfp}. These packages have been tested by a number of groups and
are well regarded as the standard tools. While the standard tools have passed
many robustness tests, they are known to be computationally intensive. The exact
amount of time needed to analyze one event depends on the length of the event.
These community tools take about a week to analyze shorter events such as binary
black hole (BBH) mergers, whereas they can take up to a month on a computer
cluster to analyze longer events such as binary neutron star (BNS) mergers. In
the coming decade, there are planned upgrade for existing facilities and
next-generation detector such as the Einstein Telescope (ET)
\cite{Punturo:2010zz} and the Cosmic Explorer (CE)
\cite{LIGOScientific:2016wof}. These upgrades will increase the sensitivity of
the detectors and allow for the detection of more events with better
signal-to-noise ratio (SNR). The number of events that will be detected in the
coming decade is expected to grow from around a thousand per year to over a
million per year \cite{Baibhav:2019gxm}. This will put a significant strain on
the current PE tools, which are not designed to handle this many events.

% Effort going to new tools development.
In order to handle the future catalog, there are efforts from multiple groups
developing tools to speed up the PE process. This includes methods that employ
modern tools such as deep learning-based method pre-trained on a large
collection of waveforms \cite{Dax:2021tsq,Dax:2022pxd}, as well as methods that
simplify the computational challenge by leveraging our knowledge about GW
signals \cite{Islam:2022afg,Roulet:2022kot}. While these methods are promising
avenues to tackle standard GW analysis tasks in the coming decade, particularly
about events involving compact binary coalescence (CBC) in GR, they capitalize
assumptions that may not be valid for analysis involving additional physical
effects such as lensing and beyond-GR analysis.

@@

For a quantitative comparison between results obtained through our code and
other existing tools, we compute the Jensen-Shannon divergence of the
marginalized distribution between our code and \texttt{Bilby}. The
Jensen-Shannon divergence is a symmetric measure of the distance between two
probability distributions, with a value of 0 indicating identical distributions
and a value of $\ln{2} \textrm{nat}$ representing the maximum possible
divergence between two distributions.

\kw{Report JSO}
